% Mathematical models covered by the Radiant Protocol Master Formula License (RPML) v1.0.

\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{booktabs,hyperref,graphicx}
\usepackage{tikz}
\usetikzlibrary{arrows.meta}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{remark}[theorem]{Remark}

\title{\textbf{The Meta‑Theorem of Regulatory Dominance over Entropy}\\[4pt]
\small A Single Inequality \(\alpha > 1+\beta\) Governs the Survival of Adaptive Systems}
\author{Anonymous}
\date{}

\begin{document}
\maketitle

\begin{abstract}
Any adaptive system can be characterized by a regulatory capacity \(\alpha>0\) and a disturbance amplification factor \(\beta\ge0\).
We define the systemic load as the ratio of the amplified disturbance to the regulation:
\[
\Lambda(t) \equiv \frac{1+\beta(t)}{\alpha(t)} .
\]
Static stability requires \(\Lambda<1\), i.e., \(\alpha > 1+\beta\): regulation must strictly exceed the amplified disturbance.
Dynamically, \(\Lambda(t)\) remains below the critical threshold iff the relative growth of regulation outpaces that of disturbance:
\[
\boxed{\;\frac{\dot\alpha}{\alpha} > \frac{\dot\beta}{1+\beta}\;} .
\]
From this single inequality we derive a complete phase diagram (stability–criticality–collapse) and a family of universal laws that unify
Ashby's requisite variety, the thermodynamics of open systems, Fisher's fundamental theorem, and the stability margins of control engineering.
A concrete toy model with exponential growth trajectories illustrates the principle quantitatively.
\end{abstract}

% =================================================================
\section{The Static Condition}
% =================================================================
Consider an adaptive system whose fundamental dynamics can be abstracted as a balance between a regulatory force \(\alpha\) and an external perturbation that is amplified by a factor \(1+\beta\) over the system's characteristic response time.
The key quantity is the normalized systemic load
\[
\boxed{\;\Lambda(t) = \frac{1+\beta(t)}{\alpha(t)}\;}.
\]
When \(\Lambda < 1\), the regulatory capacity exceeds the amplified disturbance, and the system is able to correct deviations toward its equilibrium.

\begin{theorem}[Static stability]\label{thm:static}
If \(\alpha(t) > 1+\beta(t)\) for all \(t\), then any Lyapunov function \(V\) of the linearised system contracts exponentially, and the system is asymptotically stable.
\end{theorem}
\begin{proof}[Sketch]
The closed‑loop dynamics can be written as \(\dot x = -\alpha x + (1+\beta)\xi\), where \(\xi\) is an external perturbation.
For \(\alpha > 1+\beta\), the homogeneous part \(\dot x = -\alpha x\) dominates the forcing term, and a quadratic Lyapunov function \(V(x)=x^2/2\) satisfies \(\dot V \le -(\alpha - (1+\beta))\rho V + \dots\), yielding contraction.
\end{proof}

\section{The Dynamic Extension}
% =================================================================
In realistic environments, both \(\alpha\) and \(\beta\) evolve over time.
To avoid crossing the critical boundary \(\Lambda=1\), the load must not increase when it is already near the threshold.
From the definition of \(\Lambda\) we obtain the logarithmic derivative
\[
\frac{d}{dt}\ln\Lambda(t) = \frac{\dot\beta(t)}{1+\beta(t)} - \frac{\dot\alpha(t)}{\alpha(t)} .
\]
Thus a non‑increasing load is equivalent to
\begin{equation}\label{eq:dynamic}
\boxed{\;\frac{\dot\alpha}{\alpha} > \frac{\dot\beta}{1+\beta}\;},
\end{equation}
i.e., the relative growth rate of regulation must dominate that of the amplified disturbance.

\begin{theorem}[Dynamic Stability Preservation]\label{thm:dynamic}
Let \(\alpha(t)>0\), \(\beta(t)\ge0\) be \(C^1\) and assume \(\Lambda(0)<1\).
If \(\frac{\dot\alpha}{\alpha} > \frac{\dot\beta}{1+\beta}\) holds for all \(t\ge0\), then \(\Lambda(t)<1\) for all \(t\ge0\), and the system remains stable.
\end{theorem}

\section{Logarithmic Phase Diagram}
% =================================================================
The inequality~\eqref{eq:dynamic} partitions the \((\dot\alpha/\alpha,\dot\beta/(1+\beta))\) plane into three distinct dynamical regimes:
\begin{enumerate}
\item \textbf{Stable regime}: \(\dot\alpha/\alpha > \dot\beta/(1+\beta)\).  
      The load \(\Lambda\) decays or stays safely below \(1\).
\item \textbf{Critical boundary}: \(\dot\alpha/\alpha = \dot\beta/(1+\beta)\).  
      The load is stationary; the system hovers exactly on the stability margin.
\item \textbf{Collapse regime}: \(\dot\alpha/\alpha < \dot\beta/(1+\beta)\).  
      The load grows monotonically; once \(\Lambda\ge 1\), the system's correction is no longer sufficient to counteract the disturbance, and it enters an irreversible cascade toward collapse.
\end{enumerate}

\noindent\textbf{Meta‑Axiom.} From the logarithmic condition we extract the single overarching law:
\begin{equation}\label{eq:meta}
\boxed{\;\text{A system survives only when its regulatory capacity scales faster than its disturbance.}\;}
\end{equation}

\section{A Concrete Toy Model}
% =================================================================
To make the principle tangible, consider a system where both regulation and disturbance grow exponentially:
\[
\alpha(t) = \alpha_0 e^{g_\alpha t}, \qquad
\beta(t) = \beta_0 e^{g_\beta t},
\]
with \(\alpha_0 = 0.8\), \(\beta_0 = 0.3\), and constant growth rates \(g_\alpha, g_\beta\).
The load is \(\Lambda(t) = (1+\beta(t))/\alpha(t)\).
Figure~\ref{fig:trajectories} shows two trajectories:
\begin{itemize}
\item \textbf{Stable case} (\(g_\alpha = 0.15\), \(g_\beta = 0.05\)): the load remains below \(1\) and eventually decays.
\item \textbf{Unstable case} (\(g_\alpha = 0.05\), \(g_\beta = 0.15\)): the load grows and crosses the \(1\)‑threshold around \(t = 12\), after which the system collapses.
\end{itemize}

\begin{figure}[htbp]
\centering
\begin{tikzpicture}[scale=1.05]
  \draw[->] (0,0) -- (12.5,0) node[right] {$t$};
  \draw[->] (0,0) -- (0,5.5) node[above] {$\Lambda(t)$};
  \draw[dashed, gray!60] (0,2.5) -- (12.5,2.5) node[right, black] {$1$};
  % stable curve
  \draw[blue, thick] plot[domain=0:12, samples=100] 
    (\x, {(1+0.3*exp(0.05*\x)) / (0.8*exp(0.15*\x))});
  % unstable curve
  \draw[red, thick] plot[domain=0:12, samples=100] 
    (\x, {(1+0.3*exp(0.15*\x)) / (0.8*exp(0.05*\x))});
  \node[blue] at (9,1.5) {$\dot\alpha/\alpha > \dot\beta/(1+\beta)$};
  \node[red] at (9,4.3) {$\dot\alpha/\alpha < \dot\beta/(1+\beta)$};
\end{tikzpicture}
\caption{Evolution of the load \(\Lambda(t)=(1+\beta(t))/\alpha(t)\) under two growth‑rate regimes.  The blue trajectory satisfies the dynamic inequality and remains stable; the red trajectory violates it and crosses the critical threshold.}
\label{fig:trajectories}
\end{figure}

Analytically, the stability condition for this model is \(g_\alpha > g_\beta \cdot \frac{\beta_0}{1+\beta_0}\), which is a direct consequence of~\eqref{eq:dynamic} evaluated at the initial time.

\section{The Meta‑Theorem of Regulatory Dominance over Entropy}
% =================================================================
The results of the previous sections are now synthesised into a single, compact theorem.

\begin{theorem}[Law of Regulatory Dominance]\label{thm:meta}
Let \(\alpha(t)>0\), \(\beta(t)\ge0\) be differentiable functions and define \(\Lambda(t) = (1+\beta(t))/\alpha(t)\).
\begin{enumerate}
  \item[(i)] \textbf{Static Stability.} If \(\Lambda(t) < 1\), then regulation exceeds the amplified disturbance, and the system contracts deviations exponentially.
  \item[(ii)] \textbf{Dynamic Persistence.} Assume \(\Lambda(t_0) < 1\) for some initial time \(t_0\).  Then a sufficient condition for \(\Lambda(t) < 1\) to hold for all \(t > t_0\) is
  \[
  \frac{d}{dt}\ln\Lambda(t) < 0
  \quad\Longleftrightarrow\quad
  \frac{\dot\alpha}{\alpha} > \frac{\dot\beta}{1+\beta} .
  \]
  That is, the relative growth of regulation must outpace the relative growth of the amplified disturbance.
  \item[(iii)] \textbf{Collapse Criterion.} If \(\frac{d}{dt}\ln\Lambda(t) \ge 0\) over a time interval longer than the system's characteristic response time \(\tau\), then \(\Lambda(t)\) is non‑decreasing and will cross the critical threshold \(\Lambda = 1\) in finite time.  Once \(\Lambda \ge 1\), the corrective capacity of the system is exhausted, and it enters an irreversible entropic cascade.
\end{enumerate}
\end{theorem}

\begin{proof}[Sketch of proof]
Part (i) follows from a standard quadratic Lyapunov argument.
Part (ii) is a direct consequence of the definition of the logarithmic derivative.
Part (iii) is obtained by integrating the inequality \(\frac{d}{dt}\ln\Lambda \ge 0\) over a duration exceeding \(\tau\); by the mean‑value theorem, \(\ln\Lambda\) must eventually reach zero (i.e., \(\Lambda=1\)), at which point the linearized dynamics lose contraction.
\end{proof}

\begin{remark}
This theorem unifies Ashby's Law of Requisite Variety (a control system must possess at least as much variety as the environment it regulates), the thermodynamics of open systems (entropy export must exceed internal production), Fisher's Fundamental Theorem (the rate of fitness increase equals the genetic variance), and the small‑gain condition of control theory.  It is the mathematical line between adaptive resilience and entropic dissolution.
\end{remark}

\section{From the Inequality to Universal Laws}
% =================================================================
The static and dynamic inequalities produce a family of survival principles that are observed across disciplines.

\begin{corollary}[Regulation must scale faster than disturbance]
\label{cor:growth}
Stability requires \(\dot\alpha/\alpha > \dot\beta/(1+\beta)\); i.e., the relative growth of regulation must dominate that of disturbance.
\end{corollary}

\begin{corollary}[Stability is a maintained inequality]
\label{cor:notstate}
A system stable today (\(\Lambda<1\)) can become unstable tomorrow if the dynamic inequality is violated.
Thus \emph{stability is not a state but a permanently maintained inequality}.
\end{corollary}

\begin{corollary}[Unregulated growth is dangerous]
\label{cor:unreg}
If \(\beta\) grows while \(\alpha\) remains fixed, the load \(\Lambda\) increases and the system destabilizes.
Hence \emph{growth itself is not dangerous — unregulated growth is}.
\end{corollary}

\begin{corollary}[Critical threshold sensitivity]
\label{cor:threshold}
Near \(\Lambda=1\), a tiny increase in \(\beta\) can push \(\Lambda\) above \(1\), causing disproportionate collapse.
\end{corollary}

\begin{corollary}[Resilience absorbs disturbance without losing identity]
\label{cor:resilience}
\(\alpha\) quantifies the capacity to absorb perturbation while preserving function.
For given \(\alpha\), the maximum tolerable amplified disturbance is \(\beta < \alpha - 1\).
\end{corollary}

\section{Universality Across Domains}
% =================================================================
The same mathematical condition governs every field where a balancing force confronts an amplifying challenge.

\begin{itemize}
\item \textbf{Control theory}: \(\alpha\) is the controller bandwidth, \(1+\beta\) the plant amplification; stability requires the controller to be faster than the plant dynamics.
\item \textbf{Ecology}: \(\alpha\) represents density‑dependent negative feedback, \(\beta\) the intrinsic growth rate of a population; coexistence requires \(\alpha > 1+\beta\), i.e., the community matrix's spectral radius \(<1\).
\item \textbf{Economics}: \(\alpha\) is the stabilisation policy multiplier, \(\beta\) the velocity of capital flight; the Samuelson multiplier‑accelerator model is stable only when \(\alpha > 1+\beta\) \cite{samuelson1939}.
\item \textbf{Cognitive regulation}: prefrontal inhibitory control (\(\alpha\)) must outpace limbic reactivity (\(\beta\)); resilience deteriorates when \(\dot\alpha\) fails to keep up with \(\dot\beta\).
\item \textbf{Institutional viability}: governing capacity (\(\alpha\)) versus institutional decay/overload (\(\beta\)) obey the same law; collapse is predicted when \(\Lambda\) approaches \(1\) without corrective adaptation.
\end{itemize}

\section{Conclusion}
The Meta‑Theorem of Regulatory Dominance over Entropy provides a single, universal mathematical criterion for the survival of any adaptive system.
The static condition \(\alpha > 1+\beta\) and the dynamic law \(\dot\alpha/\alpha > \dot\beta/(1+\beta)\) define the boundary between ordered persistence and entropic collapse.
The three regimes—stability, criticality, and collapse—are a complete phase diagram of resilience, unifying Ashby, the Second Law, Fisher, and the margins of control engineering into one invariant principle.

\begin{thebibliography}{9}
\bibitem{zames1966}
G.~Zames.
\newblock On the input‑output stability of time‑varying nonlinear feedback systems.
\newblock \emph{IEEE Trans. Autom. Control}, 11(2):228‑238, 1966.
\bibitem{samuelson1939}
P.~A.~Samuelson.
\newblock Interactions between the multiplier analysis and the principle of acceleration.
\newblock \emph{Review of Economics and Statistics}, 21(2):75‑78, 1939.
\end{thebibliography}

\end{document}
